---
title: "AN597- Homework 10/23- Part II"
author: "Faye Harwell"
date: "October 18, 2017"
output: html_document
---
# Homework Part II

## Non-Transformed Data... first

#### You should probably start by downloading the dataset...
```{r}
library(curl)
Stuff<- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN597_Fall17/KamilarAndCooperData.csv")
Stuf <- read.csv(Stuff, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(Stuf)
```


#### Step 1: Fit the regression model and, using {ggplot2}, produce a scatterplot with the fitted line superimposed upon the data. Append the the fitted model equation to your plot (HINT: use the function geom_text()).

```{r}
lm(Stuf$MaxLongevity_m ~ Stuf$Brain_Size_Species_Mean)-> X
X
summary(X)
```


```{r}
library(ggplot2)
Plot <- ggplot(data = Stuf, aes(x = Brain_Size_Species_Mean, y = MaxLongevity_m))
Plot <- Plot + geom_point()
Plot <- Plot + geom_smooth(method = "lm", formula = y ~ x, colour= "purple")
Plot <- Plot + annotate("text", label= "Y= 1.218X + 248.952", x= 200, y=750, color= "black" )

Plot
```

## Step #2: Identify and interpret the point estimate of the slope (β1), as well as the outcome of the test associated with the hypotheses H0: β1 = 0; HA: β1 ≠ 0. Also, find a 90 percent CI for the slope (β1) parameter.

#### From using the lm() function, you can easily determine β1= 1.218. Thus, you can support the alternative hypothesis that HA: β1 ≠ 0, since the p-value falls below the value of alpha (0.05). 

## Step #3: Using your model, add lines for the 90 percent confidence and prediction interval bands on the plot and add a legend to differentiate between the lines.

```{r}
t <- coef(summary(X))
t <- data.frame(unlist(t))
colnames(t) <- c("Est", "SE", "t", "p")
t
```

```{r}
t$calct <- (t$Est - 0)/t$SE
t$calcp <- 2 * pt(t$calct, df = 998, lower.tail = FALSE)  # x2 because is 2-tailed test
t
```

```{r}
t$lower <- t$Est - qt(0.975, df = 998) * t$SE
t$upper <- t$Est + qt(0.975, df = 998) * t$SE
ci <- c(t$lower, t$upper)  # by hand
ci
```

```{r}
ci <- confint(X, level = 0.95)  # using the results of lm()
ci
```

```{r}
h_hat <- predict(X, newdata = data.frame(Brain_Size_Species_Mean = Stuf$Brain_Size_Species_Mean))
df <- data.frame(cbind(Stuf$Brain_Size_Species_Mean, Stuf$MaxLongevity_m , h_hat))
names(df) <- c("x", "y", "yhat")
head(df)


ci <-predict(X,newdata = data.frame(Max_Longetivity = Stuf$MaxLongevity_m),interval = "confidence", level = 0.90)

head(ci)

df <- cbind(df, ci)
names(df) <- c("x", "y", "yhat", "CIfit", "CIlwr", "CIupr")
head(df)

g <- ggplot(data = df, aes(x = x, y = y))
g <- g + geom_point(alpha = 1/2)
g <- g + geom_line(aes(x = x, y = CIfit), colour = "black")
g <- g + geom_line(aes(x = x, y = CIlwr), colour = "magenta")
g <- g + geom_line(aes(x = x, y = CIupr), colour = "purple")
g <- g + annotate("text", label= "Magenta= Lower Bound CI", x= 200, y=750, color= "black")
g <- g + annotate("text", label= "Purple= Upper Bound CI", x= 200, y=790, color= "black")
g
```

## Step #4: Produce a point estimate and associated 90 percent PI for the longevity of a species whose brain weight is 800 gm. Do you trust the model to predict observations accurately for this value of the explanatory variable? Why or why not?

```{r}
pe <- predict(X, newdata = data.frame(Brain_Size_Species_Mean = 800), interval = "prediction", level = 0.90)  # for a single value
pe
```

#### I would not trust this model to accurately predict my y-variable, since many of the data do not fit within the bounds of the 90% confident intervals. Also, I cannot fully get the predict() function to give me one answer for the above equation- so there is that as well. 

# Log Transformed Data

```{r}
Stuf$Log_Brain <- log(Stuf$Brain_Size_Species_Mean)
Stuf$Log_Longetivity <- log(Stuf$MaxLongevity_m)
plot(data = Stuf, Log_Longetivity ~ Log_Brain)

lm(Stuf$Log_Longetivity ~ Stuf$Log_Brain)-> Q
Q
summary(Q)
```

```{r}
Plot <- ggplot(data = Stuf, aes(x = Log_Brain, y = Log_Longetivity))
Plot <- Plot + geom_point()
Plot <- Plot + geom_smooth(method = "lm", formula = y ~ x, colour= "pink")
Plot <- Plot + annotate("text", label= "Y= 0.2341X + 4.8790", x= 2, y=6.5, color= "black" )

Plot
```

```{r}
h_hat <- predict(Q, newdata = data.frame(Log_Brain_Size_Species_Mean = Stuf$Log_Brain))
df <- data.frame(cbind(Stuf$Log_Brain, Stuf$Log_Longetivity , h_hat))
names(df) <- c("x", "y", "yhat")
head(df)


ci <-predict(Q,newdata = data.frame(Max_Longetivity = Stuf$Log_Longetivity),interval = "confidence", level = 0.90)

head(ci)

df <- cbind(df, ci)
names(df) <- c("x", "y", "yhat", "CIfit", "CIlwr", "CIupr")
head(df)

g <- ggplot(data = df, aes(x = x, y = y))
g <- g + geom_point(alpha = 1/2)
g <- g + geom_line(aes(x = x, y = CIfit), colour = "black")
g <- g + geom_line(aes(x = x, y = CIlwr), colour = "magenta")
g <- g + geom_line(aes(x = x, y = CIupr), colour = "purple")
g <- g + annotate("text", label= "Magenta= Lower Bound CI", x= 2, y=6.5, color= "black")
g <- g + annotate("text", label= "Purple= Upper Bound CI", x= 2, y=6.35, color= "black")
g
```
### Step 5: Looking at your two models, which do you think is better? Why?

### The transformed data is likely better, but given the nature of this data, I do not think a log transformation would be justified. In the case of longetivity, do we necessarily expect to see an exponential relationship with any given morphilogical feature?