---
title: "AN597- Homework 10/23- Part II"
author: "Faye Harwell"
date: "October 18, 2017"
output: html_document
---
# Homework Part II

#### You should probably start by downloading the dataset...
```{r}
library(curl)
Stuff<- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN597_Fall17/KamilarAndCooperData.csv")
Stuf <- read.csv(Stuff, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(Stuf)
```

#### THe following steps are going far more painful than that first step, but I suppose that is okay

#### Step 1: Fit the regression model and, using {ggplot2}, produce a scatterplot with the fitted line superimposed upon the data. Append the the fitted model equation to your plot (HINT: use the function geom_text()).

```{r}
lm(Stuf$MaxLongevity_m ~ Stuf$Brain_Size_Species_Mean)-> X
X
summary(X)
```


```{r}
library(ggplot2)
Plot <- ggplot(data = Stuf, aes(x = Brain_Size_Species_Mean, y = MaxLongevity_m))
Plot <- Plot + geom_point()
Plot <- Plot + geom_smooth(method = "lm", formula = y ~ x, colour= "purple")


Plot
```

## Step #2: Identify and interpret the point estimate of the slope (β1), as well as the outcome of the test associated with the hypotheses H0: β1 = 0; HA: β1 ≠ 0. Also, find a 90 percent CI for the slope (β1) parameter.

#### From using the lm() function, you can easily determine β1= 1.218. Thus, you can support the alternative hypothesis that HA: β1 ≠ 0, since the p-value falls below the value of alpha (0.05). 

## Step #3: Using your model, add lines for the 90 percent confidence and prediction interval bands on the plot and add a legend to differentiate between the lines.

```{r}
t <- coef(summary(X))
t <- data.frame(unlist(t))
colnames(t) <- c("Est", "SE", "t", "p")
t
```

```{r}
t$calct <- (t$Est - 0)/t$SE
t$calcp <- 2 * pt(t$calct, df = 998, lower.tail = FALSE)  # x2 because is 2-tailed test
t
```

```{r}
t$lower <- t$Est - qt(0.975, df = 998) * t$SE
t$upper <- t$Est + qt(0.975, df = 998) * t$SE
ci <- c(t$lower, t$upper)  # by hand
ci
```

```{r}
ci <- confint(X, level = 0.95)  # using the results of lm()
ci
```

## Step #4: Produce a point estimate and associated 90 percent PI for the longevity of a species whose brain weight is 800 gm. Do you trust the model to predict observations accurately for this value of the explanatory variable? Why or why not?
